Face Recognition: Too Bias, or Not Too Bias?

### 2020-04-13 cvpr_workshop_backup.tex
v2

% \usepackage{mathptmx} % assumes new font selection scheme installed
% \usepackage{lipsum}
% \usepackage{slashbox}
\usepackage{booktabs}       % professional-quality tables
%\usepackage{enumitem}
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage[font=itshape]{quoting}
\usepackage{lipsum}
% \usepackage{subfig}
%% add
\usepackage{graphics}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{comment}
\usepackage{tabularx, booktabs, ragged2e}
% \usepackage{arydshln}
% \usepackage{todonotes}
% \usepackage{ctable}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{balance}

@@

\newacronym{cs}{CS}{Cosine Similiarity}

@@

% \let\vec\mathbf

@@

\def\cvprPaperID{0002} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
We reveal critical insights into problems of bias in state-of-the-art \gls{fr} systems using a novel \gls{bfw} dataset: data balanced for gender and ethnic groups. We show variations in the optimal scoring threshold for face-pairs across different subgroups. Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups. By learning subgroup-specific thresholds, we not only mitigate problems in performance gaps but also show a notable boost in the overall performance. Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception. For the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
\end{abstract}

@@

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{figures/fig1-crop.pdf}
%     \caption{Caption}
%     \label{fig:my_label}
% \end{figure}
    \begin{figure}[t!]
        \centering
        \includegraphics[width=.9\linewidth]{figures/montage.pdf}
        \caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender. The others represent the ethnicity and gender of the race and gender, respectively. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
        \label{fig:avg-faces}
    \end{figure}

@@

    The contributions in this paper are 3-fold: (1) Built a balanced dataset as a proxy to measure verification performance per subgroup; (2) Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup balances performance; (3) Showed bias causes skewness in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds; (4) Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
    % \begin{enumerate}
    %     \item Built a balanced dataset as a proxy to measure verification performance per subgroup.
    %     \item Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balance performances.
    %     \item Showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
    %     \item Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
    % \end{enumerate}

@@

        \centering
        \caption{\textbf{\gls{bfw} compared to related datasets.} Our \gls{bfw} is balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}). Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set, while only using a single resource, VGG2. \gls{rfw}, on the other hand, supports a different task (\ie domain adaptation), and focuses on race-distribution, but not the distribution of identities.}
        \scriptsize
        \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule

@@

    \subsection{Bias in \gls{fr}}
        Problems of bias in \gls{fr} have been driven by different motivations; therefore, solutions have been posed for different matters. To name a few: data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even data exploration~\cite{muthukumar2019}. Yin~\etal proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others~\cite{yin2019feature}. Similarly, others formulated the imbalanced class problem as one-shot learning, where a \gls{gan} was trained to generate face features to augment classes with fewer samples~\cite{ding2018one}. \gls{gapf} was proposed to create fair representations of the data in a quantifiable way, allowing for the finding of a de-correlation scheme from the data without access to its statistics~\cite{huang2018generative}. Wang~\etal defined subgroups at a finer-level (\ie Chinese, Japanese, Korean), and determined the familiarity of faces inter-subgroup~\cite{wang2018they}. Genders have also been used to make subgroups (\eg for analysis of gender-based face encodings~\cite{muthukumar2019}). Most recently,~\cite{wang2018racial} proposed adapting domains to bridge the bias gap by knowledge transfer, which was supported by a novel data collection, \gls{rfw}. The release of \gls{rfw} occurred after \gls{bfw} was built - although similar in terms of demographics, \gls{rfw} uses faces from MSCeleb~\cite{guo2016ms} for testing, and assumes CASIA-Face~\cite{yi2014learning} and VGG2~\cite{Cao18} were used to train. In contrast, our \gls{bfw} assumes VGG2 as the test set. 
        % As shown in \cite{wang2018racial}, MSCeleb is highly imbalanced, primarily consisting of images of white individuals. 
        % For this, we expect bias from data, for this is the training set. 
        Furthermore, \gls{bfw} balances gender and race: \gls{rfw} splits subgroups by gender and race, while \gls{bfw} splits by gender, race, or gender and race). 
        % Thus, \gls{rfw} and \gls{bfw} complement one another. % - both add demographic for subjects with faces in renowned, large-scale \gls{fr} datasets.

@@

\vspace{1mm}
\noindent\textbf{Compiling subject list.} 
Subjects were sampled from VGG2~\cite{Cao18} - unlike others that depend on multiple sources, \gls{bfw} has fewer potential conflicts in train and test overlap with existing models. We used pre-trained ethnicity~\cite{ambekar2009name} and gender~\cite{levi2015age} classifiers to find candidates for the different subgroups.

@@

\begin{figure}[!t] 
	\centering    
 \glsunset{fpr}
  \glsunset{fnr}
	\includegraphics[width=\linewidth]{figures/detcurve-improved.pdf}
		\caption{\textbf{\gls{det} curves.} \emph{Top-left}: per gender. \emph{Top-right}: per ethnicity. \emph{Bottom}: per subgroup (\ie combined). A dashed line shows a difference by about 2$\times$ in \gls{fpr} for the same threshold $\theta_{const}$. \gls{fnr} is the match error count (closer to the bottom is better).}
		\glsreset{det}\glsreset{fpr}
\label{fig:detcurves} 
\end{figure}

@@

\vspace{1mm}
\noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per subject and without overlap across folds. Furthermore, a balance in the number of faces per was obtained by sampling twenty-five faces at random from each. Next, we generated a list of all the face pairs per subject, resulting in $\sum_{l=1}^{L}\sum_{k=1}^{K_d} {N_k \choose 2}$ positive pairs, where the number of faces of all $K_l$ subjects $N_k=25$  for each of the $L$ subgroups (Table~\ref{tab:ethnic-splits}). Next, we assigned subjects to a fold. To preserve balance across folds, we sorted subjects by the number of pairs and then started assigning to alternating folds from the one with the most samples. Note, this left no overlap in identity between folds. Later, a negative set from samples within the same subgroup randomly matched until the count met that of the positive. Finally, we doubled the total count with negative pairs from across subgroups but in the same fold.
\begin{figure}[t!] 
	\glsunset{fpr}
	\glsunset{sdm}
	\centering
	\centering
	\includegraphics[width=1\linewidth]{figures/violinplots.pdf}
		\caption{\textbf{\Gls{sdm} across subgroups.} Scores of \emph{imposters} have medians around 0.3 but with variations in upper percentiles; \emph{genuine} pairs vary in mean and spread (\eg \gls{af} has more area of overlap). A threshold varying across different subgroups yields a constant \gls{fpr}.} \label{fig:detection-model} 
		\glsreset{sdm}
\end{figure}

% \vspace{-5pt}
\subsection{Problem formulation}\label{subsec:pf} 
% \gls{lfw}~\cite{LFWTech}, a commonly used benchmark for ~\gls{fr}, reserves specific train-test face-pair lists. 
\Gls{fv} is the special case of the two-class (\ie boolean) classification. Hence, pairs are labeled as the ``same'' or ``different'' \textit{genuine} pairs (\ie \textit{match}) or \textit{imposter} (\ie \textit{mismatch}), respectively. This formulation (\ie \gls{fv}) is highly practical for applications like access control, re-identification, and surveillance. Typically, training a separate model for each unique subject is infeasible. Firstly, the computational costs compound as the number of subjects increase.  Secondly, such a scheme would require model retraining each time a new person is added. Instead, we train models to encode facial images in a feature space that captures the uniqueness of a face, to then determine the outcome based on the output of a scoring (or distance) function. Formally put:
\begin{equation}\label{eg:matcher}
    f_{boolean}(\vec{x}_i, \vec{x}_j) = d(\vec{x}_i, \vec{x}_j) \leq \theta
\end{equation}

@@

@@

\begin{figure}[t!]
	\centering    
	\includegraphics[width=.85\linewidth]{figures/confusion.pdf}
		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1) for all faces of \gls{bfw} queried against all others. Notice errors concentrate within a subgroup, consistent with the \gls{sdm} in Fig.~\ref{fig:detection-model} (\ie \gls{af} show worst performance, and is mostly confused with faces of the same demographic). This plot is evidence that while race/ethnicity may be challenging to define, the subgroups are meaningful.}}
		\label{fig:confusion} 
\end{figure}

@@

\subsection{\gls{det} analysis}
\glsunset{m}
\glsunset{f}

\begin{figure*}[h!]
    \centering
    \begin{subfigure}[t]{.3\linewidth}
    \includegraphics[width=.75\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
    \caption{VGG16~\cite{simonyan2014very}}
 \end{subfigure}
    \begin{subfigure}[t]{.27\linewidth}
    \includegraphics[width=.8\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
    \caption{ResNet50~\cite{he2016deep}}
   \end{subfigure}
    \begin{subfigure}[t]{.27\linewidth}
    \includegraphics[width=.8\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
    \caption{SENet~\cite{hu2018squeeze}}
    \end{subfigure}
    \caption{\textbf{\gls{det} curves for different CNNs}. \gls{fnr} (\%) (vertical) vs \gls{fpr}  (horizontal, log-scale) for VGG2~\cite{Cao18} models with different backbones (VGG16, Resnet50, SEnet50). Lower is better. For each plot, \gls{wm} is the top-performer, while \gls{af} is the worst. The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
\end{figure*}

@@

\begin{table}[b!]
\glsunset{tar}
\glsunset{far}
\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}. For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom). Higher is better. The proposed shows improvement in all cases.}}\label{tab:ethnicy-far} 
%  \vspace{-2mm}
\begin{center}
\scriptsize
\begin{tabular}{l c c c c c}
     \gls{far} & 0.3 & 0.1 & 0.01 & 0.001 & 0.0001\\\midrule
    \multirow{2}{.1mm}{\textbf{\gls{af}}} &0.990 & 0.867 & 0.516 & 0.470 & 0.465\\[-4pt]
        &1.000 & 0.882 & 0.524 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{am}}} &0.994 & 0.883 & 0.529 & 0.482 & 0.477\\[-4pt]
        &1.000 & 0.890 & 0.533 & 0.486 & 0.482\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bf}}} &0.991 & 0.870 & 0.524 & 0.479 & 0.473\\[-4pt]
        &1.000 & 0.879 & 0.530 & 0.484 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bm}}} &0.992 & 0.881 & 0.526 & 0.480 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.532 & 0.485 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{if}}} &0.996 & 0.881 & 0.532 & 0.486 & 0.481\\[-4pt]
        &1.000 & 0.884 & 0.534 & 0.488 & 0.484\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{im}}} &0.997 & 0.895 & 0.533 & 0.485 & 0.479\\[-4pt]
        &1.000 & 0.898 & 0.535 & 0.486 & 0.481\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wf}}} &0.988 & 0.878 & 0.517 & 0.469 & 0.464\\[-4pt]
        &1.000 & 0.894 & 0.526 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wm}}} &0.989 & 0.896 & 0.527 & 0.476 & 0.470\\[-4pt]
        &1.000 & 0.910 & 0.535 & 0.483 & 0.478\\[-1pt]
    \midrule
    \multirow{2}{3mm}{\textbf{Avg.}} &0.992 & 0.881 & 0.526 & 0.478 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.531 & 0.483 & 0.479\\[-10pt]
\end{tabular}
\end{center}
\glsreset{tar}
\glsreset{far}
%  \vspace{-3mm}

@@

% \subsubsection{Other \gls{cnn} models.}~\label{app:sec:other:models}
Variations in optimal threshold exist across models (Fig.~\ref{fig:sdm-appendix-a}). Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{\href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  Sphereface as well (Fig.~\ref{fig:detcurves}). For example, the plots generated with Spherface and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst). Thus, the dditional \gls{cnn}-based models demonstrate the same phenomena: proportional to the overall model performance, exact in which the ordering subgroups insensitivity in scores space.

@@

\begin{figure}[t!] 
	\centering    
	\includegraphics[width=.85\linewidth] {figures/human_eval.pdf}
		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}. Accuracy scores shown as bar plots. Humans are most successful at recognizing their own subgroup, with a few exceptions (\eg bottom).}}
		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.
		\label{fig:human-eval} 
% 		\vspace{-1mm}
\end{figure} 
% \vspace{-4mm}
% \vspace{-5pt}

>>>>

% \usepackage{mathptmx} % assumes new font selection scheme installed
% \usepackage{lipsum}
% \usepackage{slashbox}
\usepackage{booktabs}       % professional-quality tables
%\usepackage{enumitem}
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage[font=itshape]{quoting}
\usepackage{lipsum}
% \usepackage{subfig}
%% add
\usepackage{graphics}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{comment}
\usepackage{booktabs, ragged2e}
% \usepackage{arydshln}
% \usepackage{todonotes}
% \usepackage{ctable}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{balance}

@@

\newacronym{cs}{CS}{Cosine Similarity}

@@

% \let\vec\mathbf
% \usepackage[T1]{fontenc}
% \usepackage[latin9]{inputenc}

\def\colorModel{hsb} %You can use rgb or hsb
\usepackage{babel}
\usepackage[table]{xcolor}
\usepackage{collcell}
\def\colorModel{hsb} %You can use rgb or hsb

\usepackage{hhline}
\newcommand\ColCell[1]{
  \pgfmathparse{#1<50?1:0}  %Threshold for changing the font color into the cells
    \ifnum\pgfmathresult=0\relax\color{white}\fi
  \pgfmathsetmacro\compA{0}      %Component R or H
  \pgfmathsetmacro\compB{#1/100} %Component G or S
  \pgfmathsetmacro\compC{1}      %Component B or B
  \edef\x{\noexpand\centering\noexpand\cellcolor[\colorModel]{\compA,\compB,\compC}}\x #1
  } 
\newcolumntype{E}{>{\collectcell\ColCell}m{0.4cm}<{\endcollectcell}}  %Cell width
\newcommand*\rot{\rotatebox{90}}

@@

\def\cvprPaperID{0001} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
We reveal critical insights into problems of bias in state-of-the-art \gls{fr} systems using a novel \gls{bfw} dataset: data balanced for gender and ethnic groups. We show variations in the optimal scoring threshold for face-pairs across different subgroups. Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups. By learning subgroup-specific thresholds, we not only mitigate problems in performance gaps but also show a notable boost in the overall performance. Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception. For the \gls{bfw} database, source code, and more, visit \href{http://cvpr2020.thecvf.com/submission/main-conference/author-guidelines#policies}{github}.
% \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
\end{abstract}

@@

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{figures/fig1-crop.pdf}
%     \caption{Caption}
%     \label{fig:my_label}
% \end{figure}
    \begin{figure}[t!]
        \centering
        \includegraphics[width=.8\linewidth]{figures/montage.pdf}
        \caption{\small{\textbf{\gls{bfw}.} Average face of the different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender. The others represent the ethnicity and gender of the race and gender, respectively. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}}
        \label{fig:avg-faces}
    \end{figure}

@@

    The contributions of this work are 3-fold. (1) We built a balanced dataset as a proxy to measure verification performance per subgroup for studies of bias in \gls{fr}. (2) We revealed an unwanted bias in scores of face pairs - a bias that causes ratings to skew across demographics. For this, we showed that an adaptive threshold per subgroup balances performance (\ie the typical use of a global threshold unfavorable, which we address via optimal thresholds). (3) We surveyed humans to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}.}%19-09-08}.}
    % \begin{enumerate}
    %     \item Built a balanced dataset as a proxy to measure verification performance per subgroup.
    %     \item Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balance performances.
    %     \item Showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
    %     \item Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
    % \end{enumerate}

    \subsection{Bias in \gls{fr}}
        Problems of bias in \gls{fr} have been driven by different motivations; therefore, solutions have been posed for different matters. To name a few: data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even data exploration~\cite{muthukumar2019}. Yin~\etal proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others~\cite{yin2019feature}. Similarly, others formulated the imbalanced class problem as one-shot learning, where a \gls{gan} was trained to generate face features to augment classes with fewer samples~\cite{ding2018one}. \gls{gapf} was proposed to create fair representations of the data in a quantifiable way, allowing for the finding of a de-correlation scheme from the data without access to its statistics~\cite{huang2018generative}. Wang~\etal defined subgroups at a finer-level (\ie Chinese, Japanese, Korean), and determined the familiarity of faces inter-subgroup~\cite{wang2018they}. Genders have also been used to make subgroups (\eg for analysis of gender-based face encodings~\cite{muthukumar2019}). Most recently,~\cite{wang2018racial} proposed adapting domains to bridge the bias gap by knowledge transfer, which was supported by a novel data collection, \gls{rfw}. The release of \gls{rfw} occurred after \gls{bfw} was built - although similar in terms of demographics, \gls{rfw} uses faces from MSCeleb~\cite{guo2016ms} for testing, and assumes CASIA-Face~\cite{yi2014learning} and VGG2~\cite{Cao18} were used to train. In contrast, our \gls{bfw} assumes VGG2 as the test set. 
        % As shown in \cite{wang2018racial}, MSCeleb is highly imbalanced, primarily consisting of images of white individuals. 
        % For this, we expect bias from data, for this is the training set. 
        Furthermore, \gls{bfw} balances gender and race: \gls{rfw} splits subgroups by gender and race, while \gls{bfw} has gender, race, or both). 
        % Thus, \gls{rfw} and \gls{bfw} complement one another. % - both add demographic for subjects with faces in renowned, large-scale \gls{fr} datasets.

@@

        \centering
        \caption{\small{\textbf{\gls{bfw} compared to related datasets.} Our \gls{bfw} is balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}). Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set, while only using a single resource, VGG2. \gls{rfw}, on the other hand, supports a different task (\ie domain adaptation), and focuses on race-distribution, but not the distribution of identities.}}
        \scriptsize
        \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule

@@

@@

\vspace{1mm}
\noindent\textbf{Compiling subject list.} 
Subjects were sampled from VGG2~\cite{Cao18} - unlike others built from multiple sources, \gls{bfw} has fewer potential conflicts in train and test overlap with existing models. We used pre-trained ethnicity~\cite{ambekar2009name} and gender~\cite{levi2015age} classifiers to find candidates for the different subgroups.

@@

@@

\begin{figure}[!t] 
	\centering    
 \glsunset{fpr}
  \glsunset{fnr}
	\includegraphics[width=.9\linewidth]{figures/detcurve-improved.pdf}
		\caption{\small{\textbf{\gls{det} curves.} \emph{Top-left}: per gender. \emph{Top-right}: per ethnicity. \emph{Bottom}: per subgroup (\ie combined). Dashed line shows about 2$\times$ difference in \gls{fpr} for the same threshold $\theta_{const}$. \gls{fnr} is the match error count (closer to the bottom is better).}}
		\glsreset{det}\glsreset{fpr}
\label{fig:detcurves} 
\end{figure} 
\vspace{1mm}
\noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per subject and without overlap across folds. Furthermore, a balance in the number of faces per was obtained by sampling twenty-five faces at random from each. Next, we generated a list of all the face pairs per subject, resulting in $\sum_{l=1}^{L}\sum_{k=1}^{K_d} {N_k \choose 2}$ positive pairs, where the number of faces of all $K_l$ subjects $N_k=25$  for each of the $L$ subgroups (Table~\ref{tab:ethnic-splits}). Next, we assigned subjects to a fold. To preserve balance across folds, we sorted subjects by the number of pairs and then started assigning to alternating folds from the one with the most samples. Note, this left no overlap in identity between folds. Later, a negative set from samples within the same subgroup randomly matched until the count met that of the positive. Finally, we doubled the number with negative pairs from across subgroups but in the same fold.

% \vspace{-5pt}
\subsection{Problem formulation}\label{subsec:pf} 
% \gls{lfw}~\cite{LFWTech}, a commonly used benchmark for ~\gls{fr}, reserves specific train-test face-pair lists. 
\Gls{fv} is the special case of the two-class (\ie boolean) classification. Hence, pairs are labeled as the ``same'' or ``different'' \textit{genuine} pairs (\ie \textit{match}) or \textit{imposter} (\ie \textit{mismatch}), respectively. This formulation (\ie \gls{fv}) is highly practical for applications like access control, re-identification, and surveillance. Typically, training a separate model for each unique subject is infeasible. Firstly, the computational costs compound as the number of subjects increase.  Secondly, such a scheme would require model retraining each time a new person is added. Instead, we train models to encode facial images in a feature space that captures the uniqueness of a face, to then determine the outcome based on the output of a scoring (or distance) function. Formally put:
\begin{equation}\label{eg:matcher}
    f_{boolean}(\vec{x}_i, \vec{x}_j) = d(\vec{x}_i, \vec{x}_j) \leq \theta,
\end{equation}

@@

\begin{figure}[t!] 
	\glsunset{fpr}
	\glsunset{sdm}
	\centering
	\centering
	\includegraphics[width=1\linewidth]{figures/violinplots.pdf}
		\caption{\small{\textbf{\Gls{sdm} across subgroups.} Scores of \emph{imposters} have medians around 0.3 but with variations in upper percentiles; \emph{genuine} pairs vary in mean and spread (\eg \gls{af} has more area of overlap). A threshold varying across different subgroups yields a constant \gls{fpr}.}} \label{fig:detection-model} 
		\glsreset{sdm}
\end{figure}

@@

\begin{figure}[t!]
	\centering    
	\includegraphics[width=.75\linewidth]{figures/confusion.pdf}
		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1, \%) for all faces of \gls{bfw} versus all others. Error concentrates within a subgroup - consistent with the \gls{sdm} (Fig.~\ref{fig:detection-model}), as \gls{af} performs the worst, confusing mostly within subgroup. This plot is evidence that while race/ethnicity may be challenging to define, the subgroups are meaningful.}}
		\label{fig:confusion} 
\end{figure}

@@

\begin{figure}[t!]

    % \begin{subfigure}[t]{\linewidth}

       \centering
%     \includegraphics[width=.85\linewidth]{figures/global_threshold_fpr_percent_diff.pdf}
%     \caption{Global threshold ($t_g$)}
%  \end{subfigure}
%     \begin{subfigure}[t]{\linewidth}
%       \centering
    \includegraphics[width=.95\linewidth]{figures/adaptive_fpr_percent_diff_complete.pdf}
    % \caption{Optimal threshold ($t_o$)}
%   \end{subfigure}
    \caption{\small{\textbf{Percent difference from intended \gls{fpr}.} \emph{Top:} $t_g$ yields \gls{fpr} the span as large as 2$\times$ (\ie 200\%) that intended (\ie \gls{wm} for 1e-4). Furthermore, \gls{f} subgroups tend to perform worse than intended for all cases (while \gls{m} tend to overshoot intended performance, with exception of \gls{im} in for \gls{fpr}=1e-4). \emph{Bottom:} Subgroup-specific thresholds reduces this difference to near zero, where there are small differences, the percent difference across different subgroups is fair (\ie FPR=1e-4).}}\label{fig:percent:difference}
\end{figure}
\subsection{\gls{det} analysis}
\glsunset{m}
\glsunset{f}

\begin{figure*}[h!]
    \centering
    \begin{subfigure}[t]{.3\linewidth}
    \includegraphics[width=.71\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
    \caption{VGG16~\cite{simonyan2014very}}
 \end{subfigure}
    \begin{subfigure}[t]{.27\linewidth}
    \includegraphics[width=.75\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
    \caption{ResNet50~\cite{he2016deep}}
   \end{subfigure}
    \begin{subfigure}[t]{.27\linewidth}
    \includegraphics[width=.75\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
    \caption{SENet~\cite{hu2018squeeze}}
    \end{subfigure}
    \caption{\small{\textbf{\gls{det} curves for different CNNs}. \gls{fnr} (\%) (vertical) vs \gls{fpr}  (horizontal, log-scale) for VGG2~\cite{Cao18} models with different backbones (VGG16, Resnet50, SEnet50). Lower is better. For each plot, \gls{wm} is the top-performer, while \gls{af} is the worst. The the ordering of the curves is roughly the same for each backbone.}}\label{fig:sdm-appendix-a}
\end{figure*}

@@

\begin{table}[b!]
\glsunset{tar}
\glsunset{far}
\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}. gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom). Higher is better.}}\label{tab:ethnicy-far} 
%  \vspace{-2mm}
\begin{center}
\scriptsize
\begin{tabular}{l c c c c c}
     \gls{far} & 0.3 & 0.1 & 0.01 & 0.001 & 0.0001\\\midrule
    \multirow{2}{.1mm}{\textbf{\gls{af}}} &0.990 & 0.867 & 0.516 & 0.470 & 0.465\\[-4pt]
        &1.000 & 0.882 & 0.524 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{am}}} &0.994 & 0.883 & 0.529 & 0.482 & 0.477\\[-4pt]
        &1.000 & 0.890 & 0.533 & 0.486 & 0.482\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bf}}} &0.991 & 0.870 & 0.524 & 0.479 & 0.473\\[-4pt]
        &1.000 & 0.879 & 0.530 & 0.484 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bm}}} &0.992 & 0.881 & 0.526 & 0.480 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.532 & 0.485 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{if}}} &0.996 & 0.881 & 0.532 & 0.486 & 0.481\\[-4pt]
        &1.000 & 0.884 & 0.534 & 0.488 & 0.484\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{im}}} &0.997 & 0.895 & 0.533 & 0.485 & 0.479\\[-4pt]
        &1.000 & 0.898 & 0.535 & 0.486 & 0.481\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wf}}} &0.988 & 0.878 & 0.517 & 0.469 & 0.464\\[-4pt]
        &1.000 & 0.894 & 0.526 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wm}}} &0.989 & 0.896 & 0.527 & 0.476 & 0.470\\[-4pt]
        &1.000 & 0.910 & 0.535 & 0.483 & 0.478\\[-1pt]
    \midrule
    \multirow{2}{3mm}{\textbf{Avg.}} &0.992 & 0.881 & 0.526 & 0.478 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.531 & 0.483 & 0.479\\[-10pt]
\end{tabular}
\end{center}
\glsreset{tar}
\glsreset{far}
%  \vspace{-3mm}

@@

% \subsubsection{Other \gls{cnn} models.}~\label{app:sec:other:models}
Variations in optimal threshold exist across models (Fig.~\ref{fig:sdm-appendix-a}). Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{\href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  Sphereface as well (Fig.~\ref{fig:detcurves}). For example, the plots generated with Spherface and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst). Thus, the additional \gls{cnn}-based models demonstrate the same phenomena: proportional to the overall model performance, exact in which the ordering subgroups insensitivity in scores space.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t!]
\begin{center}
    \caption{\small{\textbf{Human assessment (quantitative).} Subgroups listed per row (\ie human) and column (\ie image). Note, most do the best intra-subgroup (\textcolor{blue}{blue}), and second-best (\textcolor{red}{red}) intra-subgroup but inter-gender. WF performs the best; WF pairs are most correctly matched.}}
    % CF shows the least variation, but with the lowest accuracy. WF shows the best accuracy, in human performance and on face pairs of that subgroup.}}
    \label{tab:humsn-eval-results} 
    % \vspace{-5mm}
    %  \vspace{-2mm}
\footnotesize
\scalebox{0.94}{
\begin{tabular}{c}
\begin{tabular}{c l c c c cc}
&&\multicolumn{4}{c}{\textbf{Image}}\\
   \multicolumn{2}{c}{(Acc, \%)}& CF  & CM & WF &WM& Avg\\
\end{tabular}\\
\begin{tabular}{c l| r r r r| r}
\cline{3-7}
       &CF &\  \textbf{\textcolor{blue}{52.9}}&   \textbf{\textcolor{red}{48.0}}&43.8&44.7 &47.4 \\
         \multirow{\items}{*}{\rotatebox{90}{\textbf{Human}}} \hspace{-5mm}
        &CM &  \textbf{\textcolor{red}{45.6}} & \textbf{\textcolor{blue}{50.4}}  & 44.4 &36.2 &44.1 \\
